A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2020-05-13 manuscript.Rmd
improve code comments

```{bash, eval=FALSE}
git init # to initialize git in the current directory
git add ./data/iris.csv ./R/analysis.R # track specific files
git commit -m "add data & analysis" # take snapshot with comment
# once script or data were changed, take a new snapshot
git commit -a -m "complete data colections" # add (-a) and commit all changes
```

@@

```{bash, eval = FALSE}
git log # inspect all changes
git checkout 77db06f78e # revert local directory to previous version with hash code '77db06f78e'
```

@@

```{bash, eval=FALSE}
# link remote github repository to local directory
git remote add origin https://github.com/aaronpeikert/reproducible-research.git
# push all changes from local repository to the remote repository
git push -u origin master
```

@@

```{bash, eval=FALSE}
git clone https://github.com/aaronpeikert/reproducible-research.git
```

@@

```{bash, eval=FALSE}
report.pdf: data/iris-processed.csv report.Rmd
  Rscript -e 'rmarkdown::render("report.Rmd")'

@@

```{bash, eval=FALSE}
# obtain a local copy from the remote repository
git clone https://github.com/aaronpeikert/reproducible-research.git
# enter the local copy
cd iris
# create all files from that project
make all
```

@@

```{r,eval=FALSE}
set.seed(1234)
sample(1:10, 5)
```

@@

```{bash, eval=FALSE}
# Define the R version to be installed from rocker project
FROM rocker/verse:3.6.1
# install CRAN R packages: pacman, here, and pander
RUN install2.r --error --skipinstalled\
  pacman here pander
# install additional R packages from github: papaja and wordcountaddin
RUN installGithub.r\
  crsh/papaja benmarwick/wordcountaddin
# set the working directory inside the container
WORKDIR /home/rstudio
```

@@

```{bash, eval=FALSE}
# execute a given command in a given container image; do not save
# the state of the container; accept inputs from and return outputs to terminal
docker run --rm -it <IMAGENAME> <COMMAND>
```

@@

```{bash, eval=FALSE}
# start an interactive R session in the container named 'reproducible-research'
docker run --rm -it reproducible-research R
```

@@

```{bash, eval=FALSE}
project := $(notdir $(CURDIR))
current_dir := $(CURDIR)
home_dir := $(current_dir)
uid = --user $(shell id -u)

@@

Windows systems typically require the biggest efforts to install all necessary software. Note that you have to have Windows Pro, Enterprise, Education or Server installed, as Microsoft prevents the use of Docker on Windows Home. There is a package manager for Windows called Chocolatey, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install), providing all dependencies.
Having installed Chocolatey (and restarted the computer), all dependencies can be installed in an **admin terminal** (Windows key, then type `cmd`, right-click *Run as administrator*) via:

```{bash, eval=FALSE}
# install docker using choco
choco install -y git make docker-desktop
```

@@

```{bash, eval=FALSE}
# install docker via brew
brew cask docker
```

@@

```{bash, eval=FALSE}
# install docker via advanced package tool
apt install git make docker 
```

@@

An alternative to Docker on Linux is Singularity [@kurtzerSingularityScientificContainers2017]. Just replace any docker calls with `singularity docker`, because singularity fully supports docker images. A possible advantage is that singularity works well in high-performance computing environments and on old Linux versions, the downside is that Singularity is only available on Linux.

@@

```{r, eval=FALSE}
# BAD because of path that is specific to the computer
iris <- read.csv("/home/aaron/reproducible-research/data/iris.csv")
# GOOD because it is a relative path but slash depends on OS
iris <- read.csv("data/iris.csv")
# BETTER because truly compatible across OS
iris <- read.csv(here("data", "iris.csv"))
```

@@

```{bash, eval=FALSE}
# (1) download the complete repository containing the workflow to your computer
git clone https://github.com/aaronpeikert/workflow-showcase.git
# (2) enter the directory containing the cloned workflow
cd workflow-showcase
# (3) building the docker container
make build
# (4) running the analysis and producing the final PDF inside the container
make all DOCKER=TRUE
```

@@

```{bash, eval=FALSE}
git clone https://github.com/aaronpeikert/reproducible-research.git
cd reproducible-research
make build
make all DOCKER=TRUE
```

>>>>

```{bash, eval=FALSE}
# -- type this in a terminal --
git init # to initialize git in the current directory
git add ./data/iris.csv ./R/analysis.R # track specific files
git commit -m "add data & analysis" # take snapshot with comment
# once script or data were changed, take a new snapshot
git commit -a -m "complete data colections" # add (-a) and commit all changes
```

@@

```{bash, eval = FALSE}
# -- type this in a terminal --
git log # inspect all changes
git checkout 77db06f78e # revert local directory to previous version with hash code '77db06f78e'
```

@@

```{bash, eval=FALSE}
# -- type this in a terminal --

# link remote github repository to local directory
git remote add origin https://github.com/aaronpeikert/reproducible-research.git
# push all changes from local repository to the remote repository
git push -u origin master
```

@@

```{bash, eval=FALSE}
# -- type this in a terminal --
git clone https://github.com/aaronpeikert/reproducible-research.git
```

@@

```{bash, eval=FALSE}
# -- this is a Makefile --

report.pdf: data/iris-processed.csv report.Rmd
  Rscript -e 'rmarkdown::render("report.Rmd")'

@@

```{bash, eval=FALSE}
# -- type this in a terminal --

# obtain a local copy from the remote repository
git clone https://github.com/aaronpeikert/reproducible-research.git
# enter the local copy
cd iris
# create all files from that project
make all
```

@@

```{r,eval=FALSE}
# -- R code --
set.seed(1234)
sample(1:10, 5)
```

@@

```{bash, eval=FALSE}
# -- this is a Dockerfile --

# Define the R version to be installed from rocker project
FROM rocker/verse:3.6.1
# install CRAN R packages: pacman, here, and pander
RUN install2.r --error --skipinstalled\
  pacman here pander
# install additional R packages from github: papaja and wordcountaddin
RUN installGithub.r\
  crsh/papaja benmarwick/wordcountaddin
# set the working directory inside the container
WORKDIR /home/rstudio
```

@@

```{bash, eval=FALSE}
# -- type this in a terminal --

# execute a given command in a given container image; do not save
# the state of the container; accept inputs from and return outputs to terminal
docker run --rm -it <IMAGENAME> <COMMAND>
```

@@

```{bash, eval=FALSE}
# -- type this in a terminal --

# start an interactive R session in the container named 'reproducible-research'
docker run --rm -it reproducible-research R
```

@@

```{bash, eval=FALSE}
# -- this is a Makefile --

project := $(notdir $(CURDIR))
current_dir := $(CURDIR)
home_dir := $(current_dir)
uid = --user $(shell id -u)

@@

Windows systems typically require the biggest efforts to install all necessary pieces of software. Note that you have to have either Windows Pro, Enterprise, Education, or Server installed, as Microsoft prevents the use of Docker on Windows Home. There is a package manager for Windows called Chocolatey, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install). Chocolatey provides all software packages needed for our workflow in one place..
Having installed Chocolatey (and restarted the computer), all dependencies can be installed in an **admin terminal** (Windows key, then type `cmd`, right-click *Run as administrator*) via:

```{bash, eval=FALSE}
# -- type this in a terminal --

# install docker, make, and git using choco
choco install -y git make docker-desktop
```

@@

```{bash, eval=FALSE}
# -- type this in a terminal --

# install docker via brew
brew cask docker
```

@@

```{bash, eval=FALSE}
# -- type this in a terminal --

# install docker via advanced package tool
apt install git make docker 
```

@@

An alternative to Docker on Linux is Singularity [@kurtzerSingularityScientificContainers2017]. Just replace any docker calls with `singularity docker`, because singularity fully supports docker images. A possible advantage is that singularity works well in high-performance computing environments and on old Linux versions, the downside is that Singularity is currently only available on Linux.

@@

```{r, eval=FALSE}
# -- R code --

# BAD because of path that is specific to the computer
iris <- read.csv("/home/aaron/reproducible-research/data/iris.csv")
# GOOD because it is a relative path but slash depends on OS
iris <- read.csv("data/iris.csv")
# BETTER because truly compatible across OS
iris <- read.csv(here("data", "iris.csv"))
```

@@

```{bash, eval=FALSE}
# -- type this in a terminal --

# (1) obtain a local copy of the remote repository
git clone https://github.com/aaronpeikert/workflow-showcase.git
# (2) enter the directory containing the project
cd workflow-showcase
# (3) build the docker container
make build
# (4) run the analysis and produce the final PDF inside the container
make all DOCKER=TRUE
```

@@

```{bash, eval=FALSE}
# -- type this in a terminal --

# (1) obtain a local copy of the remote repository
git clone https://github.com/aaronpeikert/reproducible-research.git
# (2) enter the directory containing the project
cd reproducible-research
# (3) build the docker container
make build
# (4) run the analysis and produce the final PDF inside the container
make all DOCKER=TRUE
```
